---
title: "R bugzilla review"
author: "Piyush Kumar"
advisors: "Lluís Revilla, Heather Turner"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
    prettydoc::html_pretty:
        theme: cayman
        toc: true
        toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<style>
body {
    text-align: justify;
}
</style>

# Introduction

<font size="4">

## What is the bugRzilla Package?

BugRzilla is an R package that helps the user to interact with the **Bugzilla** through an API.

<font size="3">To learn more, see [bugRzilla](https://github.com/llrs/bugRzilla).</font>

## About the bugRzilla Google Summer of COde Project:-

bugRzilla is a package to interact with a bugzilla API and specially with R bugzilla. The goal of the project is to help users to submit issues to R Bugzilla.

## About the This Project:-

Explore the issues and bugs on the R Bugzilla to make the submission from bugRzilla better. It might help to identify useful patterns for R core or report the status of the R Bugzilla.
</font>

<font size="3">To learn more, see [bugzilla_viz](https://github.com/llrs/bugzilla_viz).</font>

# Setup Database on your local system

## Download SQL and MySQL Workbench

<font size="4">
To install SQL on Ubuntu one can refer a blog post by [digitalocean](https://www.digitalocean.com/community/tutorials/how-to-install-mysql-on-ubuntu-20-04).
To install MySQL workbench on Ubuntu one can refer a blog post by [linuxhint](https://linuxhint.com/installing_mysql_workbench_ubuntu/)
</font>

## Download R_bugzilla data

<font size ="4">
<ol>
<li>
The R_bugzilla data can be downloaded from  [link](https://bugs.r-project.org/db/R-bugs.sql.xz).
</li>
<li>
Since the downloaded data is a zip file so make sure you unzip the file by directly using `extract here` option to the folder you desire before dumping the file
which will have an extension `.sql` (eg: R-bugs.sql).
</li>
</ol>
</font>

## Dump downloaded R_bugzilla to MySQL workbench.

<font size="4">
Before one import the R_bugzilla SQL file one needs to create the (empty) database from MySQL if it doesn't exist already and the exported SQL don't contain CREATE DATABASE (exported with --no-create-db or -n option), before you can import it.


After considering this open your Terminal and run the command: `mysqldump -u my_username -p database_name > output_file_path`
or you can use mysql using the command: `source <Path>/R-bugs.sql;`


The options in use are:
<ol>
    <li>
        The `-u` flag indicates that the MySQL `username` will follow.
    </li>
    <li>
        The `-p` flag indicates we should be prompted for the `password` associated with the above username.
    `database_name` is of course the exact name of the database to export. eg. `bugRzilla` is the empty database you created.
    </li>
    <li>
        The `>` symbol is a Unix directive for `STDOUT`, which allows Unix commands to output the text results of the issued command to another location. In this case, that output location is a file path, specified by `output_file_path`.
    </li>
</ol>

For Example,
<ol>
    <li>
        At the command prompt, run the following command to launch the mysql shell and enter it as the root user:
        `mysql -u root -p`
    </li>
    <li>
        When you’re prompted for a password, enter the one that you set at installation time, or if you haven’t set one, press Enter to submit no password.
        The following mysql shell prompt should appear:
        `mysql>`
    </li>
    <li>
        In MySQL, I used this to dump the data in the empty database: 
        <ul>
            <li>
                Create an empty database: `create database bugRzilla;`
            </li>
            <li>
                To check whether the database is created or not use: `show databases;`
            </li>
            <li>
                Once an empty database is created then to dump the SQL data in the database use: `source /home/data/Documents/GSOC/R-bugs.sql;`
            </li>
            <li>
                To check your database is dumped correctly use: `show tables;`
```{r, eval=F}
mysql> show tables;
+---------------------+
| Tables_in_bugRzilla |
+---------------------+
| attachments         |
| bugs                |
| bugs_activity       |
| bugs_fulltext       |
| bugs_mod            |
| components          |
| longdescs           |
+---------------------+
7 rows in set (0.00 sec)
```
</li>
</ul>
</li>
</ol>
</font>

# bugRzilla Analysis

<font size="4">
For the connection to the database, I'm using the `dplyr` package, it supports to the widely-used open source databases like `MySQL`.
</font>

## The libraries used for the analysis:

```{r Loading packages}
# loading packages
library(dplyr)
library(dbplyr)
library(RMySQL)
library(DBI)
library(DT)
library(tidyverse)
library(skimr)
library(ggplot2)
library(plotly)
```

## Connect bugRzilla SQL Database with R

```{r Database Connection}
# Connecting R with MySQL
con <- dbConnect(
    MySQL(),
    dbname='bugRzilla', # change the database name to your database name
    username='root', # change the username to your username
    password='1204', # update your password
    host='localhost',
    port=3306)

#  Accessing Tables names from the Database
DBI::dbListTables(con)
```

## Data Exploartion of Bugs Table from the Database
```{r bugs table connection}
bugs_df <- tbl(con, "bugs")
#for quick view of the datatypes and the structure of data
skim(bugs_df)
```
<font size="4">
From the above table we can conclude that the few of the columns are having wrong datatype like:
<ol>
    <li>
        creation_ts
    </li>
    <li>
        delta_ts
    </li>
    <li>
        lastdiffed
    </li>
    <li>
        estimated_time
    </li>
    <li>
        remaining_time
    </li>
    <li>
        deadline
    </li>
</ol>
**Note**:The Column `estimated_time` and `remaining_time` only contains the integer value. So, It can't be transformed to Date format datatype.
Also there are columns which are empty so they are of no use of the analysis like:
<ol>
    <li>
        target_milestone
    </li>
    <li>
        qa_contact
    </li>
    <li>
        status_whiteboard
    </li>
</ol>
</font>

```{r Converting bugs table data into dataframe}
# Converting `bugs_df` to `dataframe`
bugs_df <- as.data.frame(bugs_df)
```

### Cleaning the data

<font size="4">
    First steps, check the data and prepare it for what we want:
</font>

```{r bugs_df data cleaning}
#converting the required fields in the correct datatype format
bugs_df <- bugs_df %>%
    mutate_at(vars("creation_ts", "delta_ts", "lastdiffed", "deadline"), as.Date)
# Taking the columns which are useful
bugs_df <- bugs_df %>%
    select("bug_id", "bug_severity", "bug_status", "creation_ts", "delta_ts", "op_sys", "priority", "resolution", "component_id", "version", "lastdiffed", "deadline")
#for quick view of the datatypes and the structure of data
skim(bugs_df)
#showing the `datatable`
datatable(head(bugs_df, 5), options = list(scrollX = TRUE))
```

### About the Bugs Data used for Analysis

<font size="4">
I've taken the 12 columns under consideration to Analyse the Data. The brief description about the columns as follows:
<ol>
    <li>
        **bug_id: **Unique numeric identifier for bug.
    </li>
    <li>
        **bug_severity: **How severe the bug is, e.g. enhancement, critical, etc.
    </li>
    <li>
        **bug_status: **Current status, e.g. NEW, RESOLVED, etc.
    </li>
    <li>
        **creation_ts: **When bug was filed.
    </li>
    <li>
        **delta_ts: **The timestamp of the last update on the bug. This includes updates to some related tables (e.g. "longdescs").
    </li>
    <li>
        **op_sys: **Operating system bug was seen on, e.g. Windows Vista, Linux, etc.
    </li>
    <li>
        **priority: **The priority of the bug (P1 = most urgent, P5 = least urgent).
    </li>
    <li>
        **resolution: **The resolution, if the bug is in a closed state, e.g. FIXED, DUPLICATE, etc.
    </li>
    <li>
        **component_id: **Numeric ids of the components.
    </li>
    <li>
        **version: **Version of software in which bug is seen.
    </li>
    <li>
        **lastdiffed: **The time at which information about this bug changing was last emailed to the cc list.
    </li>
    <li>
        **deadline: **Date by which bug must be fixed.
    </li>
</ol>
</font>

### Visualizations

```{r Bug created graph}
bug_created <- bugs_df %>%
    ggplot(aes(x = creation_ts, y = bug_id)) +
    geom_line(color = "darkorchid4") +
    labs(title = "Bug Creation",
           subtitle = "The data frame is sent to the plot using pipes",
           y = "Bug ID",
           x = "Date") +
    theme_bw(base_size = 15)
ggplotly(bug_created)
```
<font size="4">
    The `Time-series graph` shows that which bug_id was filed in which `month` and `year` and from the `bar graph` we can conclude that in which `year` the most bugs are filed and when one will zoom the graphs, one can see on which date which bug was filed. The most of the Bugs are filled in the month of `January` and `July`.
</font>
```{r Bug last modified graph}
last_modified <- bugs_df %>%
    ggplot(aes(x = lastdiffed, y = bug_id)) +
    geom_line() +
    labs(title = "Bug Last Modified",
         y = "Bug ID",
         x = "Date") +
    theme_bw(base_size = 15)
ggplotly(last_modified)
```

<font size="4">
    The `Time-series graph` shows that which bug_id was the last update. Most of the bugs are last updated in the month of `January`, `March`, `April`, and `July` and in the year from `2014` to `2016` most bugs are modified and in `2019` to `2020` most bugs are filed.
</font>

```{r Bug changing was last emailed to the cc list graph}
# Plotting the Time Series graph with the bug_id and delta_ts
last_modified_graph <- bugs_df %>%
    ggplot(aes(x = delta_ts, y = bug_id)) +
    geom_point() +
    labs(title = "Bug changing was last emailed to the cc list",
           y = "Bug ID",
           x = "Date") + theme_bw(base_size = 15)
ggplotly(last_modified_graph)
```

<font size="4">
    The `Time-series graph` shows that which bug_id was the last update. Most of the bugs are last updated in the month of `January`,`March`, `April`, and `July`.
</font>

```{r Bug Resolution Bar graph}
Resolution_graph <- ggplot(bugs_df,aes(x = resolution)) +
    geom_bar() +
    scale_x_discrete(guide = guide_axis(n.dodge = 5)) +
    labs(
        title = "Bug Resolution Bar graph with Bug Count",
        x = "Resolution",
        y = "Bug Count"
    ) + coord_flip()
ggplotly(Resolution_graph)
```

<font size="4">
    The `Resolution bar-graph` shows which bug_id belongs to which resolution category, if the bug is in a closed state, e.g. FIXED, DUPLICATE, etc. As we can conclude, that most bugs belong to the fixed category of the resolution.
</font>

```{r Bug Status Bar graph}
Status_graph <- ggplot(bugs_df,aes(x = bug_status)) +
    geom_bar() +
    scale_x_discrete(guide = guide_axis(n.dodge = 5)) +
    labs(
        title = "Bug Status Bar graph with Bug Count",
        x = "Bug Status",
        y = "Bug Count"
    )
ggplotly(Status_graph)
```
<font size="4">
    The `bug_status bar-graph` shows which bug_id belongs to which category of bug_status, e.g. NEW, RESOLVED, etc. As we can conclude, that most bugs belong to the closed category of the bug_status.
</font>

```{r Bug Severity Bar graph}
Severity_graph <- ggplot(bugs_df,aes(x = bug_severity)) +
    geom_bar() +
    scale_x_discrete(guide = guide_axis(n.dodge = 5)) +
    labs(
        title = "Bug Severity Bar graph with Bug Count",
        x = "Bug Severity",
        y = "Bug Count"
    )
ggplotly(Severity_graph)
```

<font size="4">
    The `bug_severity bar-graph` shows which bug_id belongs to which category of bug_severity. Most of the bugs which are filed are normal, some of the bugs that are filled under enhancements are retested for some minor and major features, and a very few bugs are filed under the blocker category.
</font>

## Data Exploartion of `bugs` and `Attachments` Table from the Database

```{r attachments table connection}
bugs_attach_df <- tbl(con, "attachments")
# Converting `bugs_attach_df` to `dataframe`
bugs_attach_df <- as.data.frame(bugs_attach_df)
#for quick view of the datatypes and the structure of data
skim(bugs_attach_df)
```
### Cleaning attachments Data

```{r attachments data cleaning}
bugs_attach_df <- bugs_attach_df %>%
    mutate_at(vars("creation_ts", "modification_time"), as.Date) %>%
    mutate_at(vars("isobsolete", "isprivate", "ispatch"), as.logical)
```

### Joining the `bugs` and `attachments` tables

```{r joing of bugs and attachments tables }
#joining the `attachments` and `bugs` table
baa <- merge(bugs_attach_df, bugs_df, by = intersect(names(bugs_attach_df), names(bugs_df)), all = TRUE)

# Created four columns `creation_month`, `creation_year` and `lastdiffed_month`, `lastdiffed_year` to find in which month and year a bug is created and modified respectively.
baa <- baa %>%
    mutate(creation_month = format(creation_ts, "%m"), 
           creation_year = format(creation_ts, "%Y"), 
           lastdiffed_month = format(lastdiffed, "%m"), 
           lastdiffed_year = format(lastdiffed, "%Y")) %>%
    group_by(creation_month, creation_year)

#showing the `datatable`
datatable(head(baa, 5), options = list(scrollX = TRUE))
```

### About the bugs_activity and attachments Data Used for Analysis

<font size="4">
    I've taken the 15 columns under consideration to Analyse the Data. The brief description about the columns as follows:
    <ol>
        <li>
            **bug_id: **Unique numeric identifier for bug.
        </li>
        <li>
            **attach_id: **Unique numeric identifier for attachment.
        </li>
        <li>
            **creation_ts: **When bug was filed.
        </li>
        <li>
            **modification_time: **The date and time on which the attachment was last modified.
        </li>
        <li>
            **description: **Text describing the attachment.
        </li>
        <li>
            **mimetype: **Content type of the attachment like `text/plain` or `image/png`.
        </li>
        <li>
            **ispatch: **Whether attachment is a patch.
        </li>
        <li>
            **filename :**Path-less file-name of attachment.
        </li>
        <li>
            **submitter_id: **Unique numeric identifier for who submitted the bug.
        </li>
        <li>
            **isobsolete: **Whether attachment is marked obsolete.
        </li>
        <li>
            **isprivate: **`TRUE` if the attachment should be `private` and `FALSE` if the attachment should be `public`.
        </li>
        <li>
            **creation_month: **The month in which the bug is created.
        </li>
        <li>
            **creation_year: **The year in which the bug is created.
        </li>
        <li>
            **lastdiffed_month: **The month in which the bug is last modified.
        </li>
        <li>
            **lastdiffed_year: **The year in which the bug is last modified.
        </li>
    </ol>
</font>

### Visualizations

```{r Bug counts}
#Counting number of bugs per month in an year
bugs_counts <- baa %>%
    arrange(bug_id) %>%
    count(creation_year)
skim(head(bugs_counts))
```
<font size = "4">
    **Note: **Here only I've shown the overview of only the 6 rows of bugs_counts Since the whole summary of the data is very large. 
</font>

```{r 3D bug count gaph}
# 3D plot to see the number of bug counts per month in a year
bug_count_graph <- plot_ly(
    x = bugs_counts$creation_month,
    y = bugs_counts$creation_year,
    z = bugs_counts$n,
    type="scatter3d", 
    mode="markers", marker = list(size=2)) 

bug_count_graph <- bug_count_graph %>%
    layout(
        title = "Bug Counts per month in a year"
    )
bug_count_graph
```
<font size = 4>
    The `3D graph` or `surface plot` is about the number of bugs counts per month in a year. The Most number of bug count is `77` in `April 2015` and the minimum bug_count is `2`.
</font>
```{r Filtering duplicate and closed bugs rows}
#filtering the data where resolution is Duplicate
res_dupli <- baa %>%
    filter(resolution == "DUPLICATE" & bug_status == "CLOSED")
```

```{r Duplicate month graph}
# plotting graph with creation month where resolution is Duplicate
dupli_month_graph <- ggplot(res_dupli) +
    geom_bar(aes(x = creation_month)) +
    labs(
        title = "Months in which Duplicate Bugs are Filed",
        x = "Months",
        y = "Bug_Count"
    )
ggplotly(dupli_month_graph)
```
<font size = 4>
    The above bar graph is about the month in which is bugs are filed where resolution is `Duplicate`. From the graph, we can see that most were filled in the month of `August` with a bug count of `12` and the least bugs are filled in the month of `December` with a bug count of `2`. This graph is from the year `2006` to `2021`.
<font>

```{r Duplicate Year graph}
# plotting graph with creation year where resolution is Duplicate
duplicate_year <- ggplot(res_dupli) +
    geom_bar(aes(x = creation_year)) +
    labs(
        title = "Year in which Duplicate Bugs are Filed",
        x = "Year",
        y = "Bug_Count"
    )
ggplotly(duplicate_year)
```

<font size = 4>
    The above Visualization is about the year in which is bugs are filed where resolution is `Duplicate`. From the graph, we can see that the most ware filled in the year `2012` having a bug count of `11` and the least bugs are filled in the year `2007`, `2008` and `2019` having a bug count of `1`. This graph is from the year `2006` to `2021`.
<font>

```{r Filtering fixed and closed bugs rows}
#filtering the data where resolution is Fixed
res_fixed <- baa %>%
    filter(resolution == "FIXED" & bug_status == "CLOSED")
```

```{r Fixed year graph}
# plotting graph with last modified year where resolution is Fixed
fixed_year_graph <- ggplot(res_fixed) +
    geom_bar(aes(x = lastdiffed_year)) +
    labs(
        title = "Year in which fixed bugs are last modified",
        x = "Year",
        y = "Bug_Count"
    ) +
    coord_flip()
ggplotly(fixed_year_graph)
```
<font size = 4>
    The above Visualization is about the year in which is bugs are last modified where resolution is `Fixed` and their status is `closed`. From the graph, we can see that the most wast last modified in the year `2002` having a bug count of `328` and In the year, `2021` `47` bugs are `fixed` and `closed`.
<font>

```{r Fixed month graph} 
# plotting graph with creation year where resolution is Fixed
fixed_month_graph <- ggplot(res_fixed) +
    geom_bar(aes(x = lastdiffed_month)) +
    labs(
        title = "Month in which fixed bugs are last modified",
        x = "Month ",
        y = "Bug_Count"
    )
ggplotly(fixed_month_graph)
```
<font size = 4>
    The `bar graph` is about the month in which is bugs are last modified where the resolution is `Fixed`. From the graph, we can see that the most wast last modified in the month `December` having a bug count of `559` and in the month of `September` having a bug count of `228` are least modified. This graph is from the year `1998` to `2021`.
<font>

```{r Invalid bugs graph}
res_invalid <- baa %>%
    filter(resolution == "INVALID" & bug_status == "CLOSED")

invalid_graph <- ggplot(res_invalid) +
    geom_bar(aes(x = creation_month)) +
    labs(
        title = "Month in which invalid bugs are filed",
        x = "Month ",
        y = "Bug_Count"
    )
ggplotly(invalid_graph)
```

<font size = 4>
    The `bar graph` is about the month in which is bugs are last modified where `resolution` is `Invalid` and `status` is `closed`. From the graph, we can see that the most were created in the month `October` having a bug count of `625` and in the month of `February` having a bug count of `491` are created. This graph is from the year `1998` to `2021`.
<font>

```{r Invaild year graph}
invaild_year_graph <- ggplot(res_invalid) +
    geom_bar(aes(x = creation_year)) +
    labs(
        title = "Year in which INVALID Bugs are Filed",
        x = "Year",
        y = "Bug_Count"
    ) + coord_flip()
ggplotly(invaild_year_graph)
```

<font size = 4>
    The `bar graph` refers to the Creation of Invalid bugs. In the year, `1998` the total of `63` Invalid bugs are created which are least, and in the year `2013` a total of `431` bugs are filed which are most.
<font>

```{r Priority graph}
priority_graph <- baa %>%
    ggplot(aes(x = creation_year, y = bug_id)) +
    geom_point() +
    facet_wrap( ~priority) +
    labs(title = "Bugs created year with their priorities",
         y = "Bug ID",
         x = "Date") + theme_bw(base_size = 9) +
    coord_flip()
ggplotly(priority_graph)
```

<font size = 4>
    The `Lattice plot` gives insight about the bugs when they are created and under which priority they fall like from the above plot we can conclude that the majority of the bugs are filed under the `P5` which is having the `least priority`.
</font>

## Data Exploration of all the tables in the database

### Data Exploartion of bugs_mod Table from the Database

```{r bugs mod connection}
bugs_mod_df <- tbl(con, "bugs_mod")
# Converting `bugs_mod_df to `dataframe`
bugs_mod_df <- as.data.frame(bugs_mod_df)
#for quick view of the datatypes and the structure of data
skim(bugs_mod_df)
```

```{r bugs_mod data table}
#showing the baa i.e `bugs_mod_df` table in the `datatable`
datatable(head(bugs_mod_df, 5), options = list(scrollX = TRUE))
```

### Data Exploartion of longdescs Table from the Database

```{r longdescs connection}
longdescs_df <- tbl(con, "longdescs")
# Converting `longdescs_df` to `dataframe`
longdescs_df <- as.data.frame(longdescs_df)
#for quick view of the datatypes and the structure of data
skim(longdescs_df)
```

```{r longdescs datatable}
#showing the baa i.e `longdescs_df` table in the `datatable`
datatable(head(longdescs_df, 5), options = list(scrollX = TRUE))
```

### Data Exploartion of bugs_activity Table from the Database

```{r loading bugs_activity table}
bugs_act_df <- tbl(con, "bugs_activity")
# Converting `longdescs_df` to `dataframe`
bugs_act_df <- as.data.frame(bugs_act_df)
#for quick view of the datatypes and the structure of data
skim(bugs_act_df)
```

### Joing all the the data tables

```{r joining all the data tables}
#joining all the data tables
total_data <- merge(bugs_df, bugs_act_df, by = intersect(names(bugs_df), 
                                                         names(bugs_act_df)), all = TRUE) %>%
    merge(., bugs_attach_df, by = intersect(names(.), 
                                            names(bugs_attach_df)), all = TRUE) %>%
    merge(., bugs_mod_df, by = intersect(names(.), 
                                         names(bugs_mod_df)), all = TRUE)
# creating a creation_year column 
total_data$creation_year <- as.Date(cut(total_data$creation_ts, 
                                         breaks = "year"))
# creating a creation_month column
total_data$creation_month <- as.Date(cut(total_data$creation_ts, 
                                         breaks = "month"))
# creating a creation_week column
total_data$creation_week <- as.Date(cut(total_data$creation_ts, 
                                        breaks = "week", start.on.monday = FALSE))

# creating a lastdiffed_year column
total_data$lastdiffed_year <- as.Date(cut(total_data$lastdiffed, 
                                           breaks = "year"))
# creating a lastdiffed_year column
total_data$lastdiffed_month <- as.Date(cut(total_data$lastdiffed, 
                                           breaks = "month"))
# creating a lastdiffed_year column
total_data$lastdiffed_week <- as.Date(cut(total_data$lastdiffed, 
                                          breaks = "week", start.on.monday = FALSE))

# selecting required columns for the analysis
total_data <- total_data %>%
    select("bug_id", "creation_ts", "bug_severity", "bug_status", "delta_ts", "op_sys", "priority", "resolution", 
           "component_id", "version", "lastdiffed", "deadline", "attach_id", "who", "bug_when", "fieldid", "added", 
           "removed", "modification_time", "description", "mimetype", "ispatch", "filename", "submitter_id", 
           "isobsolete", "isprivate", "assigned_to", "product_id", "reporter", "creation_year", "creation_month", 
           "creation_week", "lastdiffed_year", "lastdiffed_month", "lastdiffed_week")

#counting total_creation_bug_count
total_creation_bug_count <- total_data %>%
    arrange(creation_year) %>%
    count(bug_id)

# joining total_data and total_bug_count
total_data <- merge(total_creation_bug_count, total_data, by = intersect(names(total_creation_bug_count), 
                                                         names(total_data)), all = TRUE)

#for quick view of the datatypes and the structure of data
skim(total_data)
```

```{r total_data datatable}
datatable(head(total_data, 13), options = list(scrollX = TRUE))
```
### Visualizations

```{r cre_last_year_graph}
cre_last_year_graph <- total_data %>%
    ggplot(aes(y = n)) +
    geom_line(aes(x = creation_year)) +
    geom_line(aes(x = lastdiffed_year), color="steelblue", linetype="twodash") +
    labs(
        title = "Year in which bugs are Created vs Last modified",
        x = "year",
        y = "Bug_Count"
    )

ggplotly(cre_last_year_graph)
```
<font size = 4>
    In the year `2015` and `2016`, a total of `51` bugs are created and in the year `2020` a total of `51` bugs are modified as well as in the year `2015`, `31` bugs are modified. From the year `2013` to `2016` there is a peak in the creation and modification of the bugs and from the year `2003` to `2009` the is an almost linear downfall and rising in the creation and modification of the bugs.
<font>

```{r cre_last_month_graph}
cre_last_month_graph <- total_data %>%
    ggplot(aes(y = n)) +
    geom_line(aes(x = creation_month)) +
    geom_line(aes(x = lastdiffed_month), color="steelblue", linetype="twodash") +
    labs(
        title = "Month in which are Created vs Last modified",
        x = "Month",
        y = "Bug_Count"
    )

ggplotly(cre_last_month_graph)
```

<font size = 4>
    In the month `December 2015` and `January 2016`, a total of `51` bugs are created which is a sudden increase. In the `September 2016` to `October 2016` and `October 2016` to `November 2016` there is a linear increase in the creation of the bugs. From the last_modified graph, in the `November 2020` a total of `51` bugs are last modified mostly in every month an avg of `2` bugs are modified from till `2010` but from the `2011` there is an increase in the modification of the bugs having an avg of `9` bugs.
<font>

```{r cre_last_week_graph}
cre_last_week_graph <- total_data %>%
    ggplot(aes(y = n)) +
    geom_line(aes(x = creation_week)) +
    geom_line(aes(x = lastdiffed_week), color="steelblue", linetype="twodash") +
    labs(
        title = "Week in which bugs are Created vs Last modified",
        x = "Week",
        y = "Bug_Count"
    )

ggplotly(cre_last_week_graph)
```
<font size = 4>
    In the `December 2015` and `January 2016`, a total of `51` bugs are created per week which is a sudden increase. In the `September 2016` to `October 2016` and `October 2016` to `November 2016` there is a linear increase in the creation of the bugs. From the last_modified graph, in the `November 2020` a total of `51` bugs are last modified mostly in every week an avg of `2` or `3` bugs are modified from till `2010` but from the `2011` there is an increase in the modification of the bugs having an avg of `10` to `12` bugs each week.
<font>

# Conclusion

<font size = 4>
    In this project, I've visualized the bugRzilla database creating various visualization. The major graphs are the `time-series graph` which is the relationship of `bug_id`, `creation_ts`, `delta_ts`, and `last_modified` to show when the bug was created, last modified & last email to the cc list, last modified respectively. Some time-series graphs can help us to see in which week, month, the year most bugs are created.

The `bar` graph which is the relationship between the `bug_count`, `bug_status`, `bug_severity`, `resolution` to check the status like closed, new, etc, how severe the bug is like whether the bug is critical, major, etc and the resolution of the bug-like if the bug is in a closed state, e.g. FIXED, DUPLICATE, etc.

If your Bug report is effective, then its chances to get fixed are higher. It totally depends upon how effectively one reports it. Like some values are left and that should not be null to report the bugs effectively. For example, the `Priority`, `bug severity` even the `resolution` of the bugs should be specified. From the bugs data table, we can see that the column `bug_file_loc` is mostly empty like there should be proper reporting where the bug occurred. Event `Priority` is generally set from `P1` to `P5.` `P1` as “fix the bug with the highest priority” and `P5` as ” Fix when time permits” There are many rows where the priority is not mentioned which in turns takes more time by the developer to set the priority. At last in the description of the bug, these points should be there which can create the report much effective i.e.
    <ol>
        <li>
            **Reproduce steps:** Clearly, mention the steps to reproduce the bug.
        </li>
        <li>
            **Expected result:** How the application should behave on the above-mentioned steps.
        </li>
        <li>
            **Actual result:** What is the actual result of running the above steps i.e. the bug behavior.
        </li>
    </ol>
<font>

```{r Database dissconnect}
dbDisconnect(con)
```
